---
title: "DSE_Capstone_EDA_Canfield"
author: "Ryan Canfield"
date: "2024-11-10"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)      # For data manipulation.
library(ggplot2)        # For data visualizations. 
library(RColorBrewer)   # For coloring visuals.

```

```{r}
# Reading the CSV file into a data frame
df <- read.csv("cleaned_data.csv")

# Display a preview of the data frame.
head(df)

```
# Visualizations:
## Frequency Bar Charts of Demographic Variables:
### This table will present frequencies and percentages of key demographic variables, such as country, age, gender, education, and online service preferences. 

```{r}
# Specify colors for each level 
country_colors <- c("red", "gold", "orange") 
age_colors <- c("navy", "darkorchid4", "deeppink3", "salmon")
gender_colors <- c("pink", "skyblue")     
education_colors <- c("lightgoldenrod1", "springgreen3", "steelblue2", "slateblue3")  
OnlineS_colors <- c("darkred", "lightgreen")  
salary_colors <- c("deepskyblue", "darkolivegreen", "hotpink4", "firebrick3")



par(mfrow=c(2,3))

# Plot 1: Country 
barplot(table(df$Country), 
        col = country_colors,  
        xlab = "Country", 
        ylab = "Frequency",
        main = "Customer Origins", 
        cex.names = 0.8) 

# Plot 2: Online Service Preference
barplot(table(df$Online_Service_Preference), 
        col = OnlineS_colors, 
        xlab = "Online Shopping", 
        ylab = "Frequency",
        main = "Do They Prefer Online Shopping", 
        cex.names = 0.8)

# Plot 3: Gender 
barplot(table(df$Gender), 
        col = gender_colors, 
        xlab = "Gender", 
        ylab = "Frequency",
        main = "Gender of Customer", 
        cex.names = 0.8)

# Plot 4: Age 
barplot(table(df$Age), 
        col = age_colors, 
        xlab = "Age", 
        ylab = "Frequency",
        main = "Age of Customers", 
        las = 2,        
        cex.names = 0.8)
    

# Plot 5: Annual Salary 
barplot(table(df$Annual_Salary), 
        col = salary_colors, 
        xlab = "Salary", 
        ylab = "Frequency",
        main = "Customers Salary", 
        las = 2, 
        cex.names = 0.8)


```
```{r}
par(mar=c(5, 4, 4, 2) + 2.4)  # Adjust bottom margin
# Plot 6: Education 
barplot(table(df$Education), 
        col = education_colors, 
        ylab = "Frequency",
        main = "Education of Customers", 
        las = 2,        
        cex.names = 0.8)       

```


## Pie Charts of AI Tool Usage:
### These pie charts will show the proportion of customers in high, moderate, and low AI usage categories.  

```{r}
# Calculate the percentage of each category
AI_usage_percents <- df %>%
  count(AI_Usage) %>%
  mutate(Percentage = n / sum(n) * 100)

# Create the pie chart using base R
pie(AI_usage_percents$Percentage, 
    labels = paste0(AI_usage_percents$AI_Usage, ": ", round(AI_usage_percents$Percentage, 1), "%"), 
    col = c('darkred', 'steelblue2', '#28B463', 'goldenrod1'), # Customize colors
    main = "The Proportion of the Usage Amount of AI from Customers")

```

## Scatterplot of AI Trust vs. AI Satisfaction:
### This scatter plot will visualize the relationship between the newly created AI_Trust variable and AI satisfaction levels.
```{r}
unique(df$AI_Trust)
unique(df$AI_Satisfaction)

set.seed(3)  # For reproducibility

# Function to generate random number based on category
generate_random <- function(category) {
  if (category == "Low") {
    return(round(runif(1, min = 1, max = 3), 2))  # Random number between 1 and 3
  } else if (category == "Moderate") {
    return(round(runif(1, min = 4, max = 7), 2))  # Random number between 4 and 7
  } else if (category == "High") {
    return(round(runif(1, min = 8, max = 10), 2))  # Random number between 8 and 10
  }
}

# Apply the function to the 'category' column and create a new column with random numbers
df$Random_Num_based_AITrust <- sapply(df$AI_Trust, generate_random)

# View the result
head(df)
```


```{r}
df$Index <- seq_along(df$AI_Trust)


# Now create the scatter plot
ggplot(df, aes(x = Random_Num_based_AITrust, y = Index)) +
  geom_point(aes(color = AI_Trust), size = 3) +
  labs(
    title = "Relationship Between AI Trust and AI Satisfaction",
    x = "AI Trust Level",
    y = "AI Satisfaction Count"
  ) +
  theme_minimal() +
  theme(legend.title = element_blank())


```

## Histogram of AI Usage Levels: 
### This histogram will show the distribution of AI usage levels (high, moderate, low) across the dataset. 

```{r}
# Summarize the data to get frequencies
data_summary <- df %>%
  count(Country, AI_Usage)

# Create the bar graph
ggplot(data_summary, aes(x = Country, y = n, fill = AI_Usage)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Country", y = "Frequency of AI Usage", title = "AI Usage Frequency per Country") +
  scale_fill_manual(values = c("high" = "dodgerblue2", "moderate" = "#28B463", "low" = "orange", "none" = "firebrick")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank())

  
```


```{r}
# Summarize the data to get frequencies
data_summary <- df %>%
  count(AI_Tools_Used_Chatbots, AI_Usage)

# Create the bar graph
ggplot(data_summary, aes(x = AI_Tools_Used_Chatbots, y = n, fill = AI_Usage)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Used Chatbot", y = "Frequency of AI Usage", title = "AI Sites that Used Chatbots") +
  scale_fill_manual(values = c("high" = "royalblue", "moderate" = "springgreen3", "low" = "goldenrod1", "none" = "red2")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank()) 
```


```{r}
# Summarize the data to get frequencies
data_summary <- df %>%
  count(AI_Tools_Used_Virtual_Assistant, AI_Usage)

# Create the bar graph
ggplot(data_summary, aes(x = AI_Tools_Used_Virtual_Assistant, y = n, fill = AI_Usage)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Used Virtual Assistant", y = "Frequency of AI Usage", title = "AI Sites that Used Virtual Assistants") +
  scale_fill_manual(values = c("high" = "slateblue", "moderate" = "darkgreen", "low" = "goldenrod3", "none" = "violetred")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank()) 
```

```{r}
# Summarize the data to get frequencies
data_summary <- df %>%
  count(AI_Tools_Used_Voice_Photo, AI_Usage)

# Create the bar graph
ggplot(data_summary, aes(x = AI_Tools_Used_Voice_Photo, y = n, fill = AI_Usage)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Used Voice or Photo", y = "Frequency of AI Usage", title = "AI Sites that Used Photos or Voice") +
  scale_fill_manual(values = c("high" = "turquoise2", "moderate" = "yellowgreen", "low" = "khaki", "none" = "tomato")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank()) 
```
